---
title: "stats101a_project"
author: "celinenugroho"
date: "2025-03-03"
output:
  pdf_document: default
  html_document: default
---

```{r, message=FALSE}
library(dplyr)
library(car)
```


## Data cleaning

```{r}
# Importing dataset
housing <- read.csv("housing.csv")
head(housing)
```

```{r}
# Changing categorical with 3 levels to 2 for model validity
housing <- housing %>%
  mutate(furnishingstatus = ifelse(furnishingstatus == "semi-furnished", "furnished", furnishingstatus))
```

```{r}
# Converting characters to factors
housing <- housing %>%
  mutate_if(is.character, as.factor)
```

```{r}
# Encoding furnished - 1, unfurnished - 0
housing$furnishingstatus <- relevel(housing$furnishingstatus, ref = "unfurnished")

# Encoding categorical variables as 0 and 1
housing <- model.matrix(~ . -1, data = housing)
housing <-as.data.frame(housing)
head(housing)
```

```{r}
# Removing mainroadno from dataset
housing <- housing %>%
  select(-mainroadno)
```

```{r}
# Renaming columns for readability
housing <- housing %>%
  rename(mainroad = mainroadyes,
         guestroom = guestroomyes,
         basement = basementyes,
         hotwater = hotwaterheatingyes,
         ac = airconditioningyes,
         prefarea = prefareayes,
         furnishing = furnishingstatusfurnished)
```

```{r}
m1 <- lm(price ~ area + bedrooms + bathrooms + mainroad + furnishing, data = housing)
```

```{r}
summary(m1)
```

\newpage

## Scatterplot analysis

```{r}
#SCATTER PLOT FOR ALL THE VARIABLES
pairs(housing, main = "Housing Data Scatter Plot Matrix: All Variables")
```

```{r}
#histograms
par(mfrow = c(2,3))
hist(housing$price, main = "price")
hist(housing$area, main = "area")
hist(housing$bedrooms, main = "bedrooms")
hist(housing$bathrooms, main = "bathrooms")
hist(housing$mainroad, main = "mainroad")
hist(housing$furnishing, main = "furnishing")
plot.new() 
plot.new()
plot.new()
```

```{r}
#correlation Matrix
correlation_matrix <- cor(housing)
options(width = 200)
print(correlation_matrix)
```

\newpage

## Standardized residual and added variable plots

```{r}
StanRes1 <- rstandard(m1)
par(mfrow=c(2,2))
plot(housing$area, StanRes1, ylab = "Standardized Residuals", xlab = "Area")
plot(housing$bedrooms, StanRes1, ylab = "Standardized Residuals", xlab= "Bedrooms")
plot(housing$bathrooms, StanRes1, ylab = "Standardized Residuals", xlab = "Bathrooms")
plot(housing$mainroad, StanRes1, ylab = "Standardized Residuals", xlab = "mainroad")
plot(housing$furnishing, StanRes1, ylab = "Standardized Residuals", xlab = "furnishing")

```

```{r}
library(car)
par(mfrow= c(2,2))
avPlot(m1, "area", ask = FALSE)
avPlot(m1, "bedrooms", ask = FALSE)
avPlot(m1, "bathrooms", ask = FALSE)
avPlot(m1, "mainroad", ask = FALSE)
avPlot(m1, "furnishing", ask = FALSE)

```

```{r}
par(mfrow=c(2,2))
plot(m1)

#outliers
length(which(abs(rstandard(m1)) > 2))

#leverage points
n <- nrow(housing)

p <- length(m1$coefficients) - 1
length(which(hatvalues(m1) > 2 * (p+1)/n))

#influential points
length(which(cooks.distance(m1) > 4/(n-2)))
```


\newpage

## Transformation
Box-Cox transformation is not appropriate for categorical data.

1. Box-Cox on numeric variables and predictor
```{r}
summary(with(housing, powerTransform(cbind(price, area, bedrooms, bathrooms) ~ 1)))
```


2. Box-Cox on numeric variables, inverse response plot on predictor
```{r}
summary(with(housing, powerTransform(cbind(area, bedrooms, bathrooms) ~ 1)))
```

Apply Box-Cox transform on original variables
```{r}
m2 <- lm(price ~ log(area) + I(bedrooms ^ 0.5) + I(bathrooms ^ -4.5) + mainroad + furnishing, data = housing)
```

```{r}
inverseResponsePlot(m2)
```

Since $\lambda$ = 0 produces comparably small RSS to $\lambda$ = 0.11, we choose to transform the response logarithmically.

```{r}
m3 <- lm(log(price) ~ log(area) + I(bedrooms ^ 0.5) + I(bathrooms ^ -4.5) + mainroad + furnishing, data = housing)
```

$\log(\hat{price})$ = $\beta_0$ + $\beta_1 log(area)$ + $\beta_2 bathrooms^{0.5}$ + $\beta_3 bathrooms^{-4.5}$ + $\beta_4 mainroad$ + $\beta_5 furnishing$ + e

```{r}
summary(m3)
```


```{r}
par(mfrow=c(2,2))
plot(m3)

#outliers
length(which(abs(rstandard(m3)) > 2))

#leverage points
n <- nrow(housing)

p <- length(m1$coefficients) - 1
length(which(hatvalues(m3) > 2 * (p+1)/n))

#influential points
length(which(cooks.distance(m3) > 4/(n-2)))
```

```{r}
vif(m3)
```


```{r}
mreduced <- lm(log(price) ~ log(area) + I(bedrooms ^ 0.5) + mainroad + furnishing, data = housing)
anova(mreduced, m3)
```


No variable selection is needed